---
title: "Introduction to poppychat"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quickstart}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r knitr_opts, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

## Set up

To set up a Poppy chat client, first install and attach the `poppychat` package:

```{r install, eval = FALSE}
devtools::install_github("lucy-dwr/poppychat")
```

```{r attach}
library(poppychat)
```

## Get an API key

Then, ensure that you have an API key to use Poppy. **API keys are only available to State of California employees participating in the Poppy pilot.** To get an API key:

1. [Sign into Poppy](https://customeruat.sda.state.ca.gov/) in a browser
2. Navigate to Settings in the upper-right corner
3. Go to the Account tab
4. Mint an API key

## Set your API key

Once you have a Poppy API key, store it as `POPPY_API_KEY` in your `.Renviron` file so as to be secret. **Do not publish your API key in public code repositories!**

To pull up your `.Renviron` file for editing, run the following command:

```{r open_r_environ, eval = FALSE}
usethis::edit_r_environ()
```

Add the following line to your `.Renviron` file, replacing `your_api_key` with your actual API key:

```{r set_api_key, eval = FALSE}
POPPY_API_KEY = "your_api_key"
```
## Initialize a Poppy chat client

Now you can initialize a Poppy chat client. By default, the chat will retrieve your API key from your `.Renviron`. You can also set the API key directly when initializing the client by passing it as an `api_key` argument to `poppy_chat`, but this is not recommended for security reasons if your code will be shared or published.

```{r init_default_poppy_client, eval = FALSE}
poppy_llm_default <- poppy_chat()

poppy_llm_default$chat("Hi, who are you?")
#> Hello! I'm a GenAI Digital Assistant hosted by the California Department of
#> Technology (CDT) to support State of California employees. 
#>
#> I'm here to help you with:
#> - Answering questions about state procedures, policies, and resources
#> - Searching relevant documents and providing information
#> - Assisting with work-related tasks and inquiries
#> - Providing references to official documentation when available
#>
#> How can I assist you today?
```

The default model is Anthropic Claude Sonnet 4.5. You can change the model using the `model` argument.

```{r init_custom_poppy_client, eval = FALSE}
poppy_llm_gemini_25_flash <- poppy_chat(model = "Google Gemini 2.5 Flash-Lite")
```

```{r chat_gemini_25_flash, eval = FALSE}
poppy_llm_gemini_25_flash$chat("What does the California Department of Water Resources do, briefly?")
#> The California Department of Water Resources (DWR) is responsible for 
#> managing and protecting the California's water resources. They oversee flood
#> control, water conservation, and the State Water Project, which delivers
#> water across the state.
#> 
#> You can find more information on their official website: 
#> [https://water.ca.gov/](https://water.ca.gov/)
```

```{r chat_gemini_25_flash_rephrase, eval = FALSE}
poppy_llm_gemini_25_flash$chat("Could you rephrase that, briefly and in plain language suitable for a middle school student?")
#> Imagine California is a giant garden that needs lots of water to grow. The 
#> Department of Water Resources is like the garden's caretaker! They make sure 
#> there's enough water for everyone and everything – for drinking, for farms to 
#> grow food, and to keep our rivers and lakes healthy. They also help protect
#> us from floods, like building strong walls to keep rivers in their place when
#> it rains a lot.
#> 
#> Think of them as the state's water helpers!
```

To learn more about chatting with a chat client, see `{ellmer}`.

## See models

Information about Poppy models that you can use are stored in the `poppy_models` object in this package (or run `?poppy_models` from R for the help page).

```{r view_models}
dplyr::glimpse(poppy_models)
```
